---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# uid

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/matiasandina/uid/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/matiasandina/uid/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of uid is to streamline the processing of temperature data exported from UID devices. It provides functions for cleaning, outlier detection, downsampling, and diagnostic plotting, assuming a structured directory layout for raw and processed data.

## Installation

You can install the development version of uid like so:

```{r, eval=FALSE}
devtools::install_github("matiasandina/uid")
```

## Directory Structure

The package assumes the following directory layout:

```
temperature/
├── raw_data/   # Place raw UID .CSV files here
└── data/       # Cleaned, downsampled outputs will be written here
```

## Example

```{r example, eval=FALSE}
library(uid)

# Process all raw UID CSVs from a directory
process_all_uid_files(
  raw_export_dir = "temperature/raw_data",
  output_dir = "temperature/data"
)
```

This will:

- Group files by shared base name (e.g. handling _1_of_n format)
- Remove temperature outliers
- Downsample by 1-minute intervals (by default)
- Save cleaned data and diagnostic plots

***

You can also use lower-level functions for more custom workflows:

```{r, eval=FALSE}
df <- read_raw_uid_csv("path/to/file.csv")
df_flagged <- flag_temperature_outliers(df, threshold = 1)
df_clean <- df_flagged |> dplyr::filter(!outlier_global)
df_downsampled <- downsample_temperature(df_clean)

```

## Roadmap

Planned features include:

- [ ] Activity calculation from raw signal
- [ ] Interpolation of gaps
- [ ] Detection & Duration analysis for threshold crossing events
- [ ] Example Data & Tests
- [ ] Docs
