---
title: "Understanding Bout Detection and Duration Calculation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bout_duration_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dplyr)
library(ggplot2)
library(tidyr)
library(uid)
```

## Overview

This vignette illustrates how the `quantify_temp_bouts()` function detects and summarizes bouts where temperature is above or below a threshold. We'll highlight how different parameter choices affect the outcome, particularly:

- `duration_mode = "strict"` vs `"inclusive"`
- `drop_single_point_bouts`
- `fill_undetected_groups`

We'll use a toy dataset with simulated temperature values across multiple animals.

## Simulated Example Data

In this section, we define a small, handcrafted dataset with temperature readings for four animals (rfids), collected at 1-minute intervals. This toy dataset includes:

Normal temperatures (above threshold)

Brief dips below threshold

NAs to simulate missing data

One animal with all values below threshold (C3)

One animal with all values above threshold (C4)

This dataset is small enough to visually inspect but rich enough to demonstrate how quantify_temp_bouts() behaves under different scenarios.

```{r makedata}
rfids <- c("RFID_A1", "RFID_B2", "RFID_C3", "RFID_C4")
n <- 10

df <- expand_grid(
  rfid = rfids,
  common_dt = seq.POSIXt(
    as.POSIXct("2025-01-01 00:00:00"),
    by = "1 min",
    length.out = n
  )
) %>%
  mutate(
    session_name = "sample_session",
    variable = "temperature",
    value = c(
      37, 34, 34, 37, 37, 34, 37, 37, 34, 34,       # A1
      37, NA, 34, 34, 37, 37, NA, NA, 37, 34,       # B2
      rep(34, 10),                                  # C3 (all below)
      rep(37, 10)                                   # C4 (all above)
    )
  ) %>%
  group_by(rfid)
```

## Visualizing Threshold Behavior

Something that we always forget is that all "continuous" variables are recorded at a certain interval and they are in a very real sense *discrete* (read `NA` might be present, `sampling_interval` and edge cases matter).

Before diving into bout quantification, it's helpful to visualize the raw temperature traces with a threshold line at 35°C. This helps us anticipate where bouts are expected to occur.

Look for dips below the red dashed line — these are potential bouts when direction = "below".

You'll see that RFID_A1 and RFID_B2 have scattered dips, RFID_C3 is always below, and RFID_C4 never dips.

This plot provides visual ground truth for the subsequent examples.

```{r viz}
p <- df %>%
  ggplot(aes(x = common_dt, y = value)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 35, linetype = "dashed", color = "red") +
  facet_wrap(~rfid, scales = "free_y") +
  labs(title = "Temperature Time Series with Threshold = 35")

p
```

## Example 1: `duration_mode = "strict"`


Here we compute bouts using duration_mode = "strict", which calculates bout duration as end - start.

This means that bouts made of a single timepoint have a duration of zero minutes.

This mode is useful when you want to analyze exact bout spans without assuming any time extends beyond the sampled timestamps.

Look closely at the output — you’ll notice several bouts with duration_minutes == 0.

```{r example1}
strict_bouts <- quantify_temp_bouts(
  df,
  threshold = 35,
  sampling_interval = 1,
  duration_mode = "strict"
)

strict_bouts
```

## Example 2: `duration_mode = "inclusive"`

Now we run the same bout detection, but using duration_mode = "inclusive", which adds the sampling_interval to the duration.

This assumes each temperature sample represents the entire interval that follows it.

For example, a single timepoint becomes a 1-minute bout if sampling_interval = 1.

The merged table shows a direct comparison of strict vs inclusive durations. You’ll see that every inclusive value is equal to or greater than its strict counterpart.

```{r example2}
inclusive_bouts <- quantify_temp_bouts(
  df,
  threshold = 35,
  sampling_interval = 1,
  duration_mode = "inclusive"
)

inclusive_bouts
```

```{r example2b}
merge_comp <- left_join(
  strict_bouts,
  inclusive_bouts,
  by = c("rfid", "start", "end"),
  suffix = c("_strict", "_inclusive")
)

merge_comp %>%
  select(rfid, start, end, duration_minutes_strict, duration_minutes_inclusive)
```

## Example 3: Dropping single-point bouts

Here we test drop_single_point_bouts = TRUE, which removes any bout whose duration would be zero in "strict" mode.

This is helpful if you want to ignore brief threshold crossings or noise artifacts.

Important: this does not remove animals that had no bouts — only single-point bouts.

Compare this output to the one from Example 1 to see which rows were removed.

```{r example3}
drop_1pt <- quantify_temp_bouts(
  df,
  threshold = 35,
  sampling_interval = 1,
  duration_mode = "strict",
  drop_single_point_bouts = TRUE
)

drop_1pt
```

## Example 4: Filling undetected groups

In this final example, we explore what happens when an animal has no bouts at all.

When `fill_undetected_groups = TRUE`, such animals are still included in the output, with `duration_minutes = 0`.

When `fill_undetected_groups = FALSE, they are completely excluded from the result.

This is especially useful for statistical modeling or reporting where you want to include all animals in the denominator, even if they didn’t show any bouts.

```{r example4}
filled <- quantify_temp_bouts(
  df,
  threshold = 20,  # no values below this threshold
  fill_undetected_groups = TRUE
)

filled
```

```{r example4b}
empty <- quantify_temp_bouts(
  df,
  threshold = 20,
  fill_undetected_groups = FALSE
)

empty
```

## Summary Table of Options

| Parameter                 | Description                                                      |
| ------------------------- | ---------------------------------------------------------------- |
| `duration_mode`           | `"strict" = end - start`, `"inclusive" = end - start + interval` |
| `drop_single_point_bouts` | Excludes bouts where start == end (duration = 0)                 |
| `fill_undetected_groups`  | Inserts `duration = 0` for groups with no detected bouts         |

## Recommendations

- The use of `strict` or `inclusive` mode depends on your criteria for how long a bout should last and what should be counted.
- Single-point bouts (and very very short bouts!) might be returned, so you might need to filter your data.
- Fill undetected groups for complete modeling or plotting datasets (it's often the case that we want to keep true zeros in the dataset).



