---
title: "Step-by-Step UID Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step-by-Step UID Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette walks through processing a single UID CSV file using the functions in `uid`. A small example dataset is bundled with the package, and serves to illustrate how the package works.

```{r setup}
library(uid)
```

## 1. Reading raw data

```{r read, eval = FALSE}
file_path <- "path_to_your_data.csv"
# this will read and clean names, switch the datetime to proper datetime
raw <- read_raw_uid_csv(file_path)
```

The `uid` package provides sample data that emulates the result of `read_raw_uid_csv`, and we can use it for the tutorial.

```{r sampledata}
head(uid_sample_data)
```



## 2. Cleaning and outlier removal

UID raw data export might come with outliers from wrong detections. We use a simple filter based on point-to-point difference and a threshold. This approach works well for raw data with fast sampling rate (e.g., sampling rate lesser than 1 minute) and it's implemented via `flag_temperature_outliers()`. If your data comes from a slow sampling rate (e.g., sampling rate in the tens of minutes), this approach might not be the best to remove outliers. 

```{r clean}
# jumps of more than one degree will be counted as outliers
flagged <- flag_temperature_outliers(uid_sample_data, threshold = 1)
# remove temperature outliers
clean <- dplyr::filter(flagged, !outlier_global)
```

You can visualize the flagged outliers with `plot_outliers()`. This will save the plots to specific locations. 

```{r flagged-plot, eval=FALSE}
plot_outliers(flagged, output_dir = tempdir(), filepath = file_path)
```

For the purpose of this tutorial, we will plot the results here instead of saving them to file.

```{r plot}
# example for flagged
ggplot2::ggplot(flagged, ggplot2::aes(datetime, temperature, group=rfid)) +
  ggplot2::geom_line() +
  ggplot2::geom_point(
    data = flagged |> dplyr::filter(outlier_global),
    ggplot2::aes(datetime, temperature),
    color = "red"
  ) +
  ggplot2::facet_wrap(~rfid) +
  ggplot2::scale_x_datetime(date_breaks = "6 hours", date_labels = "%H:%M")
```

## 3. Activity calculation

Activity calculation is performed using the known distance between the induction coil zones as given by the `zone` column in the raw data export. You can check the zones with

```{r zonecheck}
uid:::.zone_coords
```

Euclidean distance calculations are therefore defined by the transitions from one zone to another. For example transitioning from zone 4 to zone 1 is a movement of 10.875000 inches.  

```{r trandistance}
head(uid:::.transition_distances)
```


> 🚧 Please make sure you are using the same platforms if you plan to use `calculate_activity()`

The function `calculate_activity()` uses the `.transition_distances` as a transition dictionary for platforms with 8 zones.

```{r activity}
with_activity <- calculate_activity(clean)
```

## 4. Downsampling and interpolation

The sample data is provided with a sample interval of 20 seconds. Because the timestamp that comes with real data detections will not be the same for different `rfid`, there will not be a common timestamp across `rfid`. This may or might not be a problem for different analysis, but it is also often desired to aggregate the data on longer intervals (e.g., 1 or 5 minutes). Using `downsample_temperature` or `downsample_activity()`, we can aggregate the data with precision of 1 minute. A feature of this is that the operation is performed by `rfid` (and other desired grouping variables). As a result, will have all `rfid` sampled at the same downsampled `common_time`.

```{r downsample}
down_temp <- downsample_temperature(with_activity, n = 1, precision = "minute")
down_act  <- downsample_activity(with_activity, n = 1, precision = "minute")
merged <- dplyr::left_join(
  down_temp, down_act,
  by = c("session_name", "rfid", "common_dt", "matrix_name")
)
```

Data acquisition might generate `NA` values that survive downsampling. We included some of such gaps in the sample data.

```{r}
dplyr::filter(merged, is.na(temperature))
```


It might be desired to interpolate such `NA`s using `interpolate_gaps()`. We can set `add_flag = TRUE` to check what values were interpolated.

```{r}
interp <- merged |>
  dplyr::group_by(rfid, session_name, matrix_name) |>
  interpolate_gaps(
    max_gap = 10,
    target_cols = c("temperature", "activity_index"),
    add_flag = TRUE
  )

```


```{r}
dplyr::filter(interp, .interpolated) |> 
  dplyr::select(rfid, common_dt, temperature, .interpolated)
```

By increasing the `max_gap` parameter, we will be interpolating larger gaps.

We can visualize the results of the interpolation by slicing a portion of the dataset to make the linear interpolation more evident. 

```{r}
ggplot2::ggplot(interp |> dplyr::slice(200:300, .by = rfid), ggplot2::aes(common_dt, temperature, group=rfid, color = .interpolated)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~rfid) +
  ggplot2::scale_x_datetime(date_breaks = "30 min", date_labels = "%H:%M")+
  ggplot2::scale_color_manual(values = c("TRUE" = "red", "FALSE" = "gray20"))
```



The `interp` data frame now holds cleaned, downsampled values with short gaps interpolated. For processing multiple files automatically, see the vignette on `process_all_uid_files()` or (`help("process_all_uid_files")`).
